---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

*mvforecast*: Fit, interrogate and forecast multivariate timeseries models in R
================
This is a work-in-progress package to fit a variety of multivariate timeseries models. At present, vector exponential smoothing models are supported and can be reconciled using temporal hierarchies. Both methods are established in the forecasting field but are little used or known by ecologists.

## Installation

You can install the development version from `GitHub` using:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("nicholasjclark/mvforecast")
```

You can test the package's main functionality with the `test_mvforecast` function, which fits a multivariate exponential smoothing model to the NEON *Ixodes scapularis* survey timeseries that are used in the [Ecological Forecasting Initiative's 2021 Forecast Challenge](https://ecoforecast.org/efi-rcn-forecast-challenges/). Using `thief_vets`, the function then fits an equivalent model to the original series but also uses automatic univariate forecast models applied to all higher levels of temporal aggregation to gain information about long-term trends and seasonalities that would otherwise be missed when focusing only on the noisy original data. `thief_vets` uses these aggregate forecasts to reconcile the original forecast by relying on functions available in the [`thief` package](http://pkg.robjhyndman.com/thief/). Finally, the sum of the Continuous Rank Probability Score is calculated for both models against the out-of-sample test set, where a lower score is better.
```{r}
library(mvforecast)
test_mvforecast()
```

Reconciliation is clearly a powerful technique that can drastically improve forecasts for many types of time series. In the interest of transparency and robust software engineering, a `Docker` container has been built to ensure functions in this package can be used in future if dependencies make any drastic changes. You can install the container from `Dockerhub`. For example, using `singularity` this would read as:
```{r, eval = FALSE}
singularity pull docker://nicholasjclark/brms
singularity shell brms_latest.sif
R
```

Interested users can create their own containers to add further functionality or add additional libraries. The entire `Dockerfile` reads as:
```{r, eval = FALSE}
FROM lcolling/r-verse-base:latest

# Use clang to compile Stan
# Using the default g++ causes memory issues
RUN apt-get update \
    && apt-get install -y --no-install-recommends \
    clang

RUN apt-get install -y --no-install-recommends libudunits2-dev
RUN apt-get install -y --no-install-recommends libgdal-dev
RUN apt-get update && apt-get install -y --no-install-recommends libv8-dev

RUN apt-get clean \
    && rm -rf /var/lib/apt/lists/*

# Create a makevars file and install rstan from source
# following the instructions at https://github.com/stan-dev/rstan/wiki/Installing-RStan-on-Linux
# Install remaining necessary R packages using specific versions (latest as of February 2021)
RUN R -e "options(repos = \
  list(CRAN = 'https://mran.revolutionanalytics.com/snapshot/2021-02-01/')); \
  dotR <- file.path(Sys.getenv('HOME'), '.R'); \
  if (!file.exists(dotR)) dir.create(dotR); \
  M <- file.path(dotR, 'Makevars'); \
  if (!file.exists(M)) file.create(M); \
  cat('\nCXX14FLAGS=-O3 -march=native -mtune=native -fPIC', \
  'CXX14=clang++', \
  file = M, sep = '\n', append = TRUE); \
  install.packages('rstan', type = 'source'); \
  install.packages('remotes'); \
  install.packages('brms'); \
  install.packages('here'); \
  install.packages('tidybayes'); \
  install.packages('xfun'); \
  install.packages('mgcv'); \
  install.packages('prophet'); \
  install.packages('pbapply'); \
  install.packages('ggplot2'); \
  install.packages('viridis'); \
  install.packages('reshape'); \
  remotes::install_github('nicholasjclark/mvforecast', dependencies = TRUE); \
  remotes::install_github('asael697/varstan@5378f428cad9560dae7f6daf8f431113f19a2019', dependencies = TRUE)"

CMD [ "R" ]
```

*This project is licensed under the terms of the GNU General Public License (GNU GPLv3)*
